Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-11 paper/structureScene.tex
better intro of section 4

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usetikzlibrary{arrows.meta}
\usepackage{url}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xspace}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\newcommand*{\vs}{vs.\@\xspace}

@@

As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording. Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Obrist2010}, trends in population sizes \cite{Bardeli2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.

@@

Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
\vl{Missing transition}

@@

Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR \vl{acronym ?}, much of the work dedicated to acoustic scene modeling is applied to ASC tasks.
Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task. The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams. Unless otherwise stated, scene predictions are achieved by majority voting of predictions \ml{gregoire ?}.

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}
\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
\node (omlabel1) at (3.5, 1) {$t$};
\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
\node at (2, -0.25) {(a)};
\node (omlabel2) at (8, 1) {$t$};
\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
\node at (6.5, -0.25) {(b)};
\end{tikzpicture}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. Blue and red oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
\end{center}
\end{figure}

Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay. \ml{tu peux citer le travail publié a Nature de lewicki, qui montrent que ces formes se retrouvent par ICA de sons naturels http://www.cs.cmu.edu/~learning/talks-2002/02-10-21.lewicki.pdf}

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}
\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
\node at (0.8, -0.35) {(a)};
\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
\node at (4.9, -0.35) {(b)};
\end{tikzpicture}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\end{center}
\end{figure}

@@

Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}. Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}. \gl{debate about the kind of process underlying perception: holistic / event based perception ?}.

From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}. Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene. \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.

In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach. After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.

Each scene is then described with a set of centroids. To compute the similarities between the scene based on their extracted centroids, we propose three strategies:

@@

The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
\begin{equation}
\label{eq:kc}
\begin{split}
K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
\sigma_{i} & =\parallel c_i - c_N \parallel
\end{split}
\end{equation}

where $c_N$ is defined as the $N^{\textrm{th}}$ nearest neighbor of the centroid $c_i$, and $\parallel \cdot \parallel$ denotes the Euclidean norm.

@@

\section{Experimental Evaluation}
\label{sec:ExpEval}

@@

This assertion shall be challenged in future work by considering larger databases and more diverse application scenarios, such as pleasantness prediction \cite{acta} or ecoacoustics tasks.

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
%\usetikzlibrary{arrows.meta}
\usepackage{url}
\usepackage{color}
\usepackage{graphicx}
\usepackage{xspace}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\newcommand*{\vs}{vs.\@\xspace}

@@

As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording. Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.

@@

Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \ml{Gregoire : citations ?}.
\vl{Missing transition}

\ml{Gregoire, ce paragraph est a merger avec la section motivation: Studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}. As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}. \gl{debate about the kind of process underlying perception: holistic / event based perception ?}.

@@

Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.

Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task. The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams. Unless otherwise stated, scene predictions are achieved by majority voting of predictions \ml{gregoire ?}.

@@

\begin{figure}
\begin{center}
%\begin{tikzpicture}
%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
%\node (omlabel1) at (3.5, 1) {$t$};
%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
%\node at (2, -0.25) {(a)};
%\node (omlabel2) at (8, 1) {$t$};
%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
%\node at (6.5, -0.25) {(b)};
%\end{tikzpicture}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. Blue and red oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
\end{center}
\end{figure}

Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014} \ml{missing ref}, as shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay. \ml{tu peux citer le travail publié a Nature de lewicki, qui montrent que ces formes se retrouvent par ICA de sons naturels http://www.cs.cmu.edu/~learning/talks-2002/02-10-21.lewicki.pdf}

@@

\begin{figure}
\begin{center}
%\begin{tikzpicture}
%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
%\node at (0.8, -0.35) {(a)};
%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
%\node at (4.9, -0.35) {(b)};
%\end{tikzpicture}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\end{center}
\end{figure}

@@

%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.

\ml{gregoire: une figure avec un clustering et une annotation ?}

\gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.

\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points. }

As discussed in Section \ref{sec:motivations}, several outcomes in sound perception encourage us to evaluate the validity of an object-based representation of the scene for predicting high-level properties. Though, a perfect knowledge of which event occurred at a given time is not feasible. In this paper, we thus focus on a simple quantization scheme in order to favor genericity, implemented using a clustering approach where coherent regions of the scene are identified.

One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic} that performs some variances estimates over the whole dataset and not solely on each scene, and reproject each scene in the resulting enhanced features space. Here, the aim of this process is to better balance the impact of salient events and texture like sounds on final decision by considering several ways of computing the similarity between scenes based on their centroids :

@@

The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
\begin{equation}
\label{eq:kc}
\begin{split}
K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
\sigma_{i} & =\parallel c_i - c_N \parallel
\end{split}
\end{equation}

with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.

@@

\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}

\section{Experiments}
\label{sec:ExpEval}

blurb

@@

This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse application scenarios, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} or ecoacoustics tasks.
